3D Regression Heat Map Analysis of Population Study Data

### 2015-06-26 draft.tex
added Steffis revision

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} w.r.t. a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
It shows all combinations of two to three independent features with a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
%%
Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast density data set.
%%
We were able to derive new hypotheses about relations between breast density density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.

@@

	\item An open-source web-application that can be used with data of different application domains.

@@

As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:

@@

	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.

@@

Reproducibility of results is an epidemiological key requirement and guides all analyses steps.

@@

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
\emph{Hypothesis generation} benefits from overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many epidemiological instruments available to acquire data. %, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding specific diseases or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are gathered through questionnaires, medical examinations and laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this reduces the information space and introduces a information bias, as assumptions are modeled through the discretization.
%%

@@

Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases are more likely for older subjects.
%% TODO Shorten next sentence if needed
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Confounders have to be selected by epidemiologists specific to the investigated condition.
%%
\paragraph{The Study of Health in Pomerania (SHIP).}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

@@

Different regression types are available, we focus on the following for describing linear relationships:

@@

Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.

@@

In this section we present prior and related work trying to achieve this goal.

@@

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

\paragraph{Visual analysis of epidemiological data.}
%%
The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of visual analysis of cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We incorporate their requirements, which consist of \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we focus on a fast large scale correlation analysis incorporating many features to derive insights.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%

%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their refer to their research agenda in our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM'S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al.~\cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
%The approach was applied to a cognitive aging study, where new hypotheses could be derived.

@@

We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.

@@

(1) We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.

@@

% 	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.

@@

For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.

@@

Epidemiologists argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.

@@

\textbf{Details-on-demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
%As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
The workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
%%
\paragraph{Hypothesis-free and hypothesis-based analysis.}
%%
Early analysis sessions yielded two approaches of analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis allows to derive new insights, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

%% TODO: First example can be cut if needed
$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features with the target are analyzed through this \emph{3D Regression Heat Map}.

@@

By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals f-statistics and AIC values.

@@

 The web-server (right) is written using \href{http://nodejs.org}{\texttt{Node.js}} and hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}.

@@

The finding is subject of further analysis.

>>>>

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} w.r.t. a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
It shows all combinations of two to three independent features with a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
%%
Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast density data set.
%%
We were able to derive new hypotheses about relations between breast density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS}, are not tailored to a specific target feature.

@@

	\item An open-source web application that can be used with data of different application domains.

@@

As described by Thew et al.~\cite{Thew2009}, the epidemiological workflow is divided into these different steps:

@@

	\item Clinicians/epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.

@@

Reproducibility of results is an epidemiological key requirement and guides all analysis steps.

@@

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
\emph{Hypothesis generation} benefits from overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many epidemiological instruments available to acquire data. %, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding specific diseases or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are gathered through questionnaires, medical examinations and laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this reduces the information space and introduces an information bias, as assumptions are modeled through the discretization.
%%

@@

Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases are more likely for older subjects.
%% TODO Shorten next sentence if needed
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Confounders have to be selected by epidemiologists specific to the investigated condition.
%%
\paragraph{The Study of Health in Pomerania (SHIP).}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data.
%%
A second cohort, \texttt{SHIP-TREND}, was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and is the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

@@

Different regression types are available; we focus on the following for describing linear relationships:

@@

Tukey already stated in 1977 that data are too often analyzed solely using a confirmatory data analysis \cite{Tukey}.

@@

In this section, we present prior and related work trying to achieve this goal.

@@

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once, as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color-coding distances to the line-of-fit.
%%
The model parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

\paragraph{Visual analysis of epidemiological data.}
%%
The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of a visual analysis of cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of a cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We incorporate their requirements, which consist of a \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we focus on a fast large scale correlation analysis incorporating many features to derive insights.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on the \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.

Generalized pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%

%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high-dimensional data.
%%
Their refer to their research agenda in our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high-dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high-dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM'S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al.~\cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
%The approach was applied to a cognitive aging study, where new hypotheses could be derived.

@@

We try to reproduce their results and further investigate findings presented by Niemann et al.~\cite{Niemann2014}.

@@

(1) We focus on the large-scale analysis of a vast number of linear and logistic regression models by assessing their quality-of-fit using descriptive metrics.

@@

% 	\item We focus on the large-scale analysis of a vast number of linear and logistic regression models by assessing their quality-of-fit using descriptive metrics.

@@

For explorative analysis, they preferred an approach which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.

@@

Epidemiologists argued that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.

@@

\textbf{Details-on-demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
%As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
The workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
%%
\paragraph{Hypothesis-free and hypothesis-based analysis.}
%%
Early analysis sessions yielded two approaches of analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis allows to derive new insights, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer-associated features).
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and data set features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

%% TODO: First example can be cut if needed
$Cancer \sim X + Y + Z$ is the formulation of a hypothesis where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features with the target are analyzed through this \emph{3D Regression Heat Map}.

@@

By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals, f-statistics and AIC values.

@@

 The web server (right) is written using \href{http://nodejs.org}{\texttt{Node.js}} and hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}.

@@

The finding is subject to further analysis.
